Reproducible Research Peer Assessment 1
========================================================


## Loading and preprocessing the data.
```{r,echo=TRUE}
activity <- read.csv(unz("activity.zip", "activity.csv"), header = TRUE)
#Convert the "date" variable structure from factor to date.
activity$date <- as.Date(activity$date,format="%Y-%m-%d")
library(plyr)
steps.per.day <- ddply(activity, .(date), function(df) sum(df$steps))
# Rename second column name
colnames(steps.per.day)[2] <- "steps"
```


## What is mean total number of steps taken per day?
Histogram of the total number of steps taken each day.
```{r,echo=TRUE,fig.height=6,fig.width=8}
library(ggplot2)
qplot(steps, data = steps.per.day, xlab = "Datetime", ylab = "Total Steps",binwidth = 2000,main="Histogram of Total Steps by Day") + geom_histogram(binwidth = 2000,fill = "firebrick", color = "steelblue")
```


Calculate the mean of steps taken per day.
```{r,echo=TRUE}
mean(steps.per.day$steps,na.rm=TRUE)
```


Calculate the median of steps taken per day.
```{r,echo=TRUE}
median(steps.per.day$steps,na.rm=TRUE)
```

## What is the average daily activity pattern?
Time Series plot of mean steps by day
```{r,echo=TRUE,fig.height=6,fig.width=8}
qplot(data=activity, interval, steps, stat='summary', fun.y='mean',geom="line", xlab="Interval",ylab="Mean Steps")
```


5-minute interval with the max number of steps across all the averages
```{r,echo=TRUE}
max.interval <- ddply(activity, .(date,interval), function(df) mean(df$steps,na.rm=TRUE))
max(max.interval$V1,na.rm=TRUE)
```


Total number of missing values in the dataset
```{r,echo=TRUE}
sum(is.na(activity))
```

## Imputing missing values
Fill in all of the missing values in the dataset with the mean number of steps that are taken for the coresponding interval that had an NA.
```{r,echo=TRUE,cache=TRUE}
# Find the mean number of steps for each 5 minute interval
mean.interval <- ddply(activity, .(interval), function(df) median(df$steps,na.rm=TRUE))
# Rename second column 
colnames(mean.interval)[2] <- "steps"
# Loop to locate and replace NAs
for (i in seq_along(activity$steps)){
     if (is.na(activity[i,]$steps)){
     x <- which(activity[i,]$interval==mean.interval$interval)
     activity[i,]$steps <- mean.interval[x,]$steps
 }
}
```


Histogram of the total number of steps taken each day after NAs were filled.
```{r,echo=TRUE,fig.height=6,fig.width=8}
# Calculate mean steps per day
steps.per.day <- ddply(activity, .(date), function(df) sum(df$steps))
# Rename second column name
colnames(steps.per.day)[2] <- "steps"
qplot(steps, data = steps.per.day, xlab = "Datetime", ylab = "Total Steps",binwidth = 2000,main="Histogram of Total Steps by Day") + geom_histogram(binwidth = 2000,fill = "steelblue", color = "firebrick")
```


Recalculate the sum of steps per day now that NAs are filled in.
```{r,echo=TRUE}
steps.per.day <- ddply(activity, .(date), function(df) sum(df$steps,na.rm=TRUE))
# Rename second column name
colnames(steps.per.day)[2] <- "steps"
```

Calculate the mean of steps taken per day.
```{r,echo=TRUE}
mean(steps.per.day$steps,na.rm=TRUE)
```


Calculate the median of steps taken per day.
```{r,echo=TRUE}
median(steps.per.day$steps,na.rm=TRUE)
```

## Are there differences in activity patterns between weekdays and weekends?
Create a new variable in the dataset with indicating whether a given date is a weekday or weekend day.
```{r,echo=TRUE}
activity$week <- ifelse (weekdays(activity$date)%in%c("Saturday","Sunday"),"weekend","weekday")

# Convert the variable to a factor format.
activity$week <- as.factor(activity$week)  
```


Panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r,echo=TRUE,fig.height=6,fig.width=9}
qplot(data=activity, interval, steps, stat='summary', fun.y='mean',geom="line", xlab="Interval",ylab="Mean Steps")+ facet_grid(week ~ .)+ aes(colour = factor(week))
```
